---
title: "Shiny Workshop, Nov 13th, 2019"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---


## What is Shiny?

Shiny is an R package that makes it easy to build interactive web applications (apps) straight from R.

Let's look at some Shiny examples:
<br>
<a href="https://shiny.rstudio.com/gallery/telephones-by-region.html" target="_blank">Telephones by region</a>
<br>
<a href="https://shiny.rstudio.com/gallery/single-file-shiny-app.html" target="_blank">Random histogram</a>
<br>
<a href="https://shiny.rstudio.com/gallery/lego-set.html" target="_blank">Lego Set Database</a>

## Let's look under the hood of a Shiny App
Before doing anything, we need to load the Shiny library (click the play button to run the code). We're also going to use the tidyverse for plotting & data filtering.
```{r}
library(shiny)
library(tidyverse)
```


A Shiny app has three main components:
* a user interface (ui) object - contains everything you see
* a server function - handles user input, sends output to ui
* a call to the shinyApp function

### Layout

Below is the code for a super basic Shiny app. It doesn't really do anything, but includes some functions or how the user interface is presented:
```{r}
ui <- fluidPage(
    titlePanel("Title of my Shiny App"),
    sidebarLayout(
      sidebarPanel("I'll put user inputs here"),
      mainPanel("My outputs will go here")
    )
  
  ) # the user interface

server <- function(input, output) {
  
            #will receive the input, do things with it, and send back to ui 
            # but currently does nothing
  
          } #the server

shinyApp(ui = ui, server = server) # this launches your app
```

The ui object contains some pretty specific layout functions (fluidpage, titlePanel, sidebarLayout, sidebarPanel, mainPanel). These are just a few Shiny functions, with many more described in the documentation: https://shiny.rstudio.com/reference/shiny/1.3.2/

YOUR TURN:
Below, create a shiny app with  ui, server, and shinyApp components. Label the parts different from above.
```{r}


```


### User Inputs

Shiny allows for a variety of inputs: text input, numeric input, select menus, checkboxes, datepickers, etc. A complete list is available here: https://shiny.rstudio.com/tutorial/written-tutorial/lesson3/

For now, we're going to add a select menu, so the user can select a value from a pulldown menu:

```{r}
ui <- fluidPage(
    titlePanel("Title of my Shiny App"),
    sidebarLayout(
      sidebarPanel(
        selectInput("mySelectID", "Make a choice!", 
                choices = list("One" = 1, "Two" = 2, "Three" = 3), selected = 1) 
        ),
      mainPanel("My outputs will go here")
    )
  
  ) # the user interface

server <- function(input, output) {
            #will receive the input, do things with it, and send back to ui 
            # but currently does nothing
          }

shinyApp(ui = ui, server = server) # this launches your app


```

Now we're going to add some code to the server function, so it will receive the selected input, and send it back to the output.

```{r}
ui <- fluidPage(
    titlePanel("Title of my Shiny App"),
    sidebarLayout(
      sidebarPanel(
        selectInput("mySelectID", "Make a choice!", 
                choices = list("One" = 1, "Two" = 2, "Three" = 3), selected = 1) 
        ),
      mainPanel(
        textOutput("theOutput")  #matches the output ID in line 117
        )
    )
  ) # the user interface

server <- function(input, output) {
    output$theOutput <- renderText({   # use "renderText" when the output is text!
      # add any R code as needed in here
      theNumber<-input$mySelectID  # matches the select ID on line 105
      paste("You have selected", theNumber) # "paste" concatenates strings  
  })
          }

shinyApp(ui = ui, server = server) # this launches your app


```

YOUR TURN:
Create a Shiny app with a Select menu containing the values 10,20,30,40, and 50. In the server function, multiply the selected value by 5, and make the output read "<number selected> times 5 is <product>".
Hint: you'll need to convert the input from string to numeric before multiplying, using the as.numeric function (e.g. num<-as.numeric("5"))

```{r}


```

### Working with a dataset

The dplyr dataset "population" shows population data by country from 1995-2013:  

```{r}
population

```

We can make a line plot of population change over time for any country....here's France:

```{r}
ggplot(data=filter(.data=population, country=="France"), mapping=aes(x=year, y=population))+geom_line()

```

Let's say we want to build a Shiny app in which we let the user select the country, resulting in a similar plot of population change over time. First thing to do is figure out how to create a select menu from the population data set. We can do this using the "unique" function:

```{r}
countries<-unique(population$country)

#this maps 
names(countries)<-countries

countries

```




```{r}
countries<-unique(population$country)
names(countries)<-countries

ui <- fluidPage(
    titlePanel("Population over time by country"),
    sidebarLayout(
      sidebarPanel(
        selectInput("myCountry", "Select a country", choices = countries) 
        ),
      mainPanel(
        plotOutput("myPlot")  #matches the output ID in line 181
        )
    )
  ) # the user interface

server <- function(input, output) {
    output$myPlot <- renderPlot({   # use "renderPlot" when the output is a plot!
      ggplot(data=filter(.data=population, country==input$myCountry), mapping=aes(x=year, y=population))+geom_line()+labs(title=paste("population of",input$myCountry, "1995-2017"))
    })
}

shinyApp(ui = ui, server = server) # this launches your app


```


YOUR TURN:
Create a Shiny app that has the following:
* uses the babyNames data set
* user selects the year from a select menu 
* output is a bar chart showing the five most popular names

Bonus: Can you add a second user input that lets the user choose Boys or Girls?

Hints:
This loads the babyNames dataset:
babyNames<-read_csv("babyNames.csv");

This yields the top 5 for a given Year and Sex:
top5<-filter(.data=babyNames, Year=="1975" & Sex=="Girls") %>% arrange(desc(Count)) %>% head(5)

This plots a bar chart from the top5 data:
ggplot(data=top5, mapping=aes(x=Name, y=Count))+geom_bar(stat="identity")+labs(title="Most popular baby names - 1975")


```{r}


```



