---
title: "Shiny Workshop, Nov 13th, 2019"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---


### What is Shiny?

Shiny is an R package that makes it easy to build interactive web applications (apps) straight from R.

Let's look at some Shiny examples:

Alcohol Consumption by Country: https://viz.datasci.watzek.cloud/alcohol/

Lego Set Database: https://shiny.rstudio.com/gallery/lego-set.html

Bookie Clicker (by Gabe): https://viz.datasci.watzek.cloud/bookieclicker/


### Loading the Shiny package

Before doing anything, we need to load the Shiny library (click the play button to run the code). We're also going to use the tidyverse for plotting & data filtering.
```{r}
library(shiny)
library(tidyverse)
```


### Shiny App Structure

A Shiny app has three main components:
* a user interface (ui) object - contains everything you see
* a server function - handles user input, sends output to ui
* a call to the shinyApp function

Below is the code for a super basic Shiny app. It doesn't really do anything, but includes some functions or how the user interface is presented:
```{r}

####### The UI ##########

ui <- fluidPage(
    titlePanel("Title of my Shiny App"),
    sidebarLayout(
      sidebarPanel("I'll put user inputs here"),
      mainPanel("My outputs will go here")
    )
  
  ) 

####### The Server ###########

server <- function(input, output) {
            #will receive the input, do things with it, and send back to ui 
            # but currently does nothing
          }

###### call to shinyApp function ########

shinyApp(ui = ui, server = server) # this launches your app
```

The ui object contains some pretty specific layout functions (fluidpage, titlePanel, sidebarLayout, sidebarPanel, mainPanel). These are just a few Shiny functions, with many more described in the documentation: https://shiny.rstudio.com/reference/shiny/1.3.2/

YOUR TURN:
Below, create a shiny app with  ui, server, and shinyApp components. Add your own custom text to the UI.
```{r}


```


### User Inputs

Shiny allows for a variety of inputs: text input, numeric input, select menus, checkboxes, datepickers, etc. A complete list is available here: https://shiny.rstudio.com/tutorial/written-tutorial/lesson3/

For now, we're going to add a select menu, so the user can select a value from a pulldown menu:

```{r}
ui <- fluidPage(
    titlePanel("Title of my Shiny App"),
    sidebarLayout(
      sidebarPanel(
        selectInput(inputId= "mySelectID", "Make a choice!", 
                choices = list("One" = 1, "Two" = 2, "Three" = 3)) 
        ),
      mainPanel("My outputs will go here")
    )
  
  ) 

server <- function(input, output) {
            #will receive the input, do things with it, and send back to ui 
            # but currently does nothing
          }

shinyApp(ui = ui, server = server) 


```

Now we're going to add some code to the server function, so it will receive the selected input, and send it back to the output.

```{r}
ui <- fluidPage(
    titlePanel("Title of my Shiny App"),
    sidebarLayout(
      sidebarPanel(
        selectInput(inputId="mySelectID", "Make a choice!", 
                choices = list("One" = 1, "Two" = 2, "Three" = 3)) 
        ),
      mainPanel(
        textOutput(outputId="theOutput")  #matches the output$theOutput below 
        )
    )
  ) 

server <- function(input, output) {
    output$theOutput <- renderText({   # use "renderText" when the output is text!
      # add any R code as needed in here
      
      theNumber<-input$mySelectID  # matches the inputId in the selectInput function
      paste("You have selected", theNumber) # "paste" concatenates strings  
  })
          }

shinyApp(ui = ui, server = server) 


```

YOUR TURN:
Create a Shiny app with a Select menu containing the values 10,20,30,40, and 50. In the server function, multiply the selected value by 5, and make the output read "<number selected> times 5 is <product>".

Hint: you'll need to convert the input from string to numeric before multiplying, using the as.numeric function (e.g. num<-as.numeric(input$mySelectID))

```{r}


```

### Working with a dataset

The dplyr dataset "population" shows population data by country from 1995-2017:  

```{r}
population

```

We can make a line plot of population change over time for any country....here's France:

```{r}
ggplot(data=filter(.data=population, country=="France"), mapping=aes(x=year, y=population))+geom_line()

```

Let's say we want to build a Shiny app in which we let the user select the country, resulting in a similar plot of population change over time. First thing to do is figure out how to create a select menu from the population data set. We can do this using the "unique" function:

```{r}
countries<-unique(population$country)

#this creates a "list" that can be used in a select input:
# list("Afghanistan" = "Afghanistan", "Albania" = "Albania") etc.
names(countries)<-countries

countries

```

Now we can create the Shiny app. First we'll get our list ready for the select menu. Then we can add the ggplot function in the "renderPlot" function in the server, while substituting the input value from "myCountry" in the ggplot data filter and title:


```{r}
countries<-unique(population$country)
names(countries)<-countries

ui <- fluidPage(
    titlePanel("Population over time by country"),
    sidebarLayout(
      sidebarPanel(
        selectInput(inputId="myCountry", "Select a country", choices = countries) 
        ),
      mainPanel(
        plotOutput(outputId="myPlot")  #matches output$myPlot below
        )
    )
  ) # the user interface

server <- function(input, output) {
    output$myPlot <- renderPlot({   # use "renderPlot" when the output is a plot!
      
      ggplot(data=filter(.data=population, country==input$myCountry), mapping=aes(x=year, y=population))+geom_line()+labs(title=paste("population of",input$myCountry, "1995-2017"))
      
    })
}

shinyApp(ui = ui, server = server) 

```


YOUR TURN:
Create a Shiny app that has the following:
* uses the babyNames data set
* user selects the year from a select menu 
* output is a bar chart showing the five most popular names

Bonus: Can you add a second user input that lets the user choose Boys or Girls?

Hints:
This loads the babyNames dataset:
babyNames<-read_csv("babyNames.csv");

This yields the top 5 for a given Year and Sex:
top5<-filter(.data=babyNames, Year=="1975" & Sex=="Girls") %>% arrange(desc(Count)) %>% head(5)

This plots a bar chart from the top5 data:
ggplot(data=top5, mapping=aes(x=Name, y=Count))+geom_bar(stat="identity")+labs(title="Most popular baby names - 1975")


```{r}


```


### Parting thoughts

* We have a shiny server where you can deploy your shiny apps  
* Want to kick around an idea for a shiny app? Talk to us (digitalinitiatives@lclark.edu)
* Thank you!



